Reproducible Research: Course Project 1  
=======================================  
1.  Code for reading in the dataset and/or processing the data
``` {r echo=T}
activity <- read.table(file="/Users/joe/Downloads/activity.csv", header=T, sep=",")
dim(activity)
summary(activity)
```

Marking a variable indicating the day of the week for each observation
``` {r echo=T}
activity$day_of_week <- weekdays(as.Date(activity$date))
table(activity$day_of_week)
```
  
  
2. Histogram of the total number of steps taken each day  
``` {r echo=T}
date_steps <- tapply(activity$steps, activity$date, sum, na.rm=T)
print(date_steps)
hist(date_steps, xlab="Total Number of Steps", main="Total Number of Steps Each Day", col="grey", breaks=10)
```


3. Mean and median number of steps taken each day
``` {r echo=T}
mean(date_steps)
median(date_steps)
```


4. Time series plot of the average number of steps taken
```{r echo=T}
interval_steps <- tapply(activity$steps, activity$interval, mean, na.rm=T)
plot(names(interval_steps), interval_steps, type="l", xlab="Time Interval", ylab="Average Number of Steps", main="Average Number of Steps per Time Interval")
```


5. The 5-minute interval that, on average, contains the maximum number of steps
```{r echo=T}
names(interval_steps)[which(interval_steps == max(interval_steps))]
```

6. Code to describe and show a strategy for imputing missing data
```{r echo=T}
# Total number of missing values
sum(is.na(activity$steps))
# Proportion
mean(is.na(activity$steps))

# Imuputing missing number of steps per interval using Poisson regression
p_fit <- glm(steps ~ interval + day_of_week, data=activity, family=poisson)
summary(p_fit)
imputed_values <- predict(p_fit, activity, type="response")

# New dataset with imputed missing steps
activity_imputed <- activity

# Apply imputed values to NAs
activity_imputed$impute_steps <- activity_imputed$steps # non missing same as steps
activity_imputed$impute_steps[which(is.na(activity_imputed$steps))] <- imputed_values[which(is.na(activity_imputed$steps))] # missing get imputed value
# Check that imputation was applied and no NA values exist in impute_steps variable
sum(is.na(activity_imputed$impute_steps))
```

7. Histogram of the total number of steps taken each day after missing values are imputed
```{r echo=T}
# Histogram of imputed steps
date_impute_steps <- tapply(activity_imputed$impute_steps, activity_imputed$date, sum)
hist(date_impute_steps, xlab="Total Number of Steps", main="Total Number of Steps Each Day", col="grey", breaks=10)

# Mean and Median of inputed steps
mean(date_impute_steps)
median(date_impute_steps)

# Do the values differ from the estimate from the first part?
t.test(date_steps, date_impute_steps, paired=T)
wilcox.test(date_steps, date_impute_steps, paired=T, exact=F, conf.int=T)
```

Paired ttest and paired wilcoxon signed rank test show a significant difference between mean of paired difference and median of paired difference.  Both the mean and median of the non-imputed minus imputed paired differences have negative estimates, which shows that the imputated values are higher than the non-imputed values.
  
  
8.Panel plot comparing the average number of steps taken per 5-minute inteval across weekdays and weekends
```{r echo=T, fig.height=9}
activity_imputed$weekday <- factor(activity_imputed$day_of_week %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), labels=c("Weekend", "Weekday"))

table(activity_imputed$weekday)

interval_steps_weekday <- tapply(activity_imputed$steps, list(activity_imputed$interval, activity_imputed$weekday), mean, na.rm=T)


layout(matrix(c(1,2), 2, 1))
plot(names(interval_steps_weekday[,1]), interval_steps_weekday[,1], type="l", xlab="Time Interval", ylab="Average Number of Steps", main=paste(colnames(interval_steps_weekday)[1],"Average Number of Steps per Time Interval"))
plot(names(interval_steps_weekday[,2]), interval_steps_weekday[,2], type="l", xlab="Time Interval", ylab="Average Number of Steps", main=paste(colnames(interval_steps_weekday)[2],"Average Number of Steps per Time Interval"))
```
